---
title: "m254-hw1 Name: Jiahao Tian"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ISLR)
dim(NCI60$data)
head(NCI60$labs) 
table(NCI60$labs)
```
**Problem 1**

1. Order the 6, 830 p-values from the smallest to the largest. Plot the ordered p-values as the y-axis and their corresponding ranks as the x-axis. If all the 6, 830 null hypotheses are true, the p-values should be uniformly distributed. Are the observed p-values likely from a uniform distribution?

**Solution:** Yes, it is uniform distribution.

```{r}
NSCLC <- NCI60$data[NCI60$labs=='NSCLC',] 
RENAL <- NCI60$data[NCI60$labs=='RENAL',] 
p_values <- mapply(function(i) 
  t.test(NSCLC[,i], RENAL[,i], var.equal = T, 
         paired = F, alternative = "two.sided")$p.value, 
  1:ncol(NSCLC))

sort_p <- sort(p_values)
plot(sort_p, xlab = "Rank", ylab = "p value", 
     cex=0.1, col='blue', main = "Sorted p value")

plot(ecdf(sort_p), main="Empirical Dist. of Ordered p-values", col="blue")
lines(x=seq(0,1,0.01), y=seq(0,1,0.01), col="red", lwd=1.5)
text(0.8, 0.6, "Theoretical cdf", col="red")
text(0.25, 0.6, "Empirical cdf", col="blue")
```

2. What will be the p-value cutoﬀ if you would like to control the Family Wise Error Rate un-
 der 0.05 using the Bonferroni correction? How many genes will be identiﬁed as diﬀerentially
 expressed using this cutoﬀ?
 
 **Solution:** P-value cutoff will be around 3e-04, and none of the gene will 
 be identified as differentially expressed.
 
 
```{r}
 plot(sort_p, cex=0.5, xlim = c(1, 25), ylim = c(0, 0.0008),
     main = "FDR and FWER", xlab = 'index i', ylab = 'p-value', pch=8)

x <- c(1:6830); fdr <- c(x/6830*0.05); bonforroni <- rep(0.05 / 6830, times=length(x));
holm <- c(0.05/(length(x)-x+1))
sum(sort_p <= bonforroni[1])
```

