---
title: "m254-hw1 Name: Jiahao Tian"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ISLR)
dim(NCI60$data)
head(NCI60$labs) 
table(NCI60$labs)
```
**Problem 1**

1. Order the 6, 830 p-values from the smallest to the largest. Plot the ordered p-values as the y-axis and their corresponding ranks as the x-axis. If all the 6, 830 null hypotheses are true, the p-values should be uniformly distributed. Are the observed p-values likely from a uniform distribution?

**Solution:** Yes, it is uniform distribution.

2. What will be the p-value cutoﬀ if you would like to control the Family Wise Error Rate un-
 der 0.05 using the Bonferroni correction? How many genes will be identiﬁed as differentially
 expressed using this cutoﬀ?
 
3. What will be the p-value cutoﬀ if you would like to control the False Discovery Rate under 0.05
 using the Benjamini-Hochberg procedure? How many genes will be identiﬁed as diﬀerentially
 expressed using this cutoﬀ?
 
4. What will be the p-value cutoﬀ if you would like to control the Per-Comparison Error Rate
 (i.e., signiﬁcance level) under 0.05? How many genes will be identiﬁed as diﬀerentially expressed
 using this cutoﬀ?
 

```{r}
NSCLC <- NCI60$data[NCI60$labs=='NSCLC',] 
RENAL <- NCI60$data[NCI60$labs=='RENAL',] 
p_values <- mapply(function(i) 
  t.test(NSCLC[,i], RENAL[,i], var.equal = T, 
         paired = F, alternative = "two.sided")$p.value, 
  1:ncol(NSCLC))

sort_p <- sort(p_values)
plot(sort_p, xlab = "Rank", ylab = "p value", 
     cex=0.1, col='blue', main = "Sorted p value")

plot(ecdf(sort_p), main="Empirical Dist. of Ordered p-values", col="blue")
lines(x=seq(0,1,0.01), y=seq(0,1,0.01), col="red", lwd=1.5)
text(0.8, 0.6, "Theoretical cdf", col="red")
text(0.25, 0.6, "Empirical cdf", col="blue")

plot(sort_p, cex=0.5, xlim = c(1, 25), ylim = c(0, 0.0008),
     main = "FDR and FWER", xlab = 'index i', ylab = 'p-value', pch=8)

x <- c(1:6830); fdr <- c(x/6830*0.05); bonforroni <- rep(0.05 / 6830, times=length(x));
holm <- c(0.05/(length(x)-x+1))
sum(sort_p <= bonforroni[1])

lines(x, fdr,  lwd=1.5)
lines(x, bonforroni,  lwd=1.5)
lines(x, holm, lwd=0.5)
text(20, 0.00001, "Holm's")
text(23, 0.00001, "Bonforroni")
text(18, 0.0002, "FDR")

 sum(p_values<=0.05)
```
 
**Problem 2**

Use the NCI60 data, give an example to each of the following data analysis tasks (please refer to the Leek and Peng, Science (2015) articl). Note: this is an open question;diﬀerent people may give diﬀ erence examples to the same task.

1. Descriptive study
2. Exploratory study
3. Inferential study
4. Predictive study
5. Causal study
6. Mechanistic study

```{r}
## Descriptive study
table(NCI60$labs)

## Exploratory study
plot(round(apply(NCI60$data, 2, mean)[1:20], 4), xlab="Cell type", 
     ylab="Gene expression", 
     main="Exploratory of first 10 genes", pch=3)

## Inferential study
plot(NCI60$data[,1], NCI60$data[,2], xlab="Gene 1", ylab="Gene 2", cex=0.4, 
     pch=1, main = "Inferential study of gene 1 and 2")

## Predictive study
library(e1071)
dat <- data.frame(NCI60$data[,1:2], y=factor(NCI60$labs))
mod <- svm(y~., data=dat, kernel="linear", cost=10, scale = F)
plot(mod, dat)

```

**Problem 3**

library(tidyverse)

```{r}
rm(list=ls())
babies <- read.table("babies.txt", header=TRUE)
bwt.nonsmoke <- subset(babies, smoke==0)$bwt
bwt.smoke <- subset(babies, smoke==1)$bwt

#1. We will generate the following statistics based on a sample size of 10 and observe the following diﬀerence:

permutation_test <- function(N=10, method='median', B=1000, randomseed=1) {

  n <- N
  set.seed(randomseed)
  nonsmokers <- sample(bwt.nonsmoke , n)
  smokers <- sample(bwt.smoke , n)
  if(method=='mean') {
    diff <- mean(smokers) - mean(nonsmokers)
  } else if(method=='median') diff <- median(smokers) - median(nonsmokers)
  else return("Choose between 'mean' and 'median' !")
  
 if(B > 100000) return("Iteration cannot be greater than 1000000 !")
  diff_permute <- c()
  for (i in 1:B) {
    dat <- c(smokers, nonsmokers)
    shuffle <- sample(dat)
    smokers_star <- shuffle[1:n]
    nonsmokers_star <- shuffle[(n+1):(2*n)]
    if(method=='mean') {
      diff_star <- mean(smokers_star) - mean(nonsmokers_star)
    } else if(method=='median') {
      diff_star <- median(smokers_star) - median(nonsmokers_star)
    }
    diff_permute <- c(diff_permute, diff_star)
  }
  p <- (sum(diff_permute <= -abs(diff)) + sum(diff_permute >= abs(diff))) / B
  return(list(test_statstic=diff, permutation=diff_permute, p_value=p))
}
result_mean <- permutation_test(N=10, method = 'mean', B=1000, randomseed=1)
result_mean$p_value

hist(result_mean$permutation, breaks = 20,  xlab='null statistic',
     main = 'Hist. of null dist. (N=10, B=1000, mean)')
abline(v=result_mean$test_statstic, lwd=5)
text(-10, 80, paste("Test stat. with p value:", result_mean$p_value), cex=1.5)

#2. Repeat the above exercise, but instead of the diﬀerences in mean, consider the diﬀerences in median.

result_med  <- permutation_test(N=10, method = 'median', B=1000, randomseed=1)
result_med$p_value

hist(result_med$permutation, breaks = 20, xlab='null statistic',
     main = 'Hist. of null dist. (N=10, B=1000, median)')
abline(v=result_med$test_statstic, lwd=5)
text(-10, 80, paste("Test stat. with p value:", result_med$p_value), cex=1.5)
```

**Problem 4**

```{r}
#1. Please display the Q-Q plots and argue whether the Gaussian assumption is reasonable for this data set.

qqnorm(bwt.nonsmoke, main = "QQ plot for nonsmokers")
qqline(bwt.nonsmoke,col=2)

qqnorm(bwt.smoke, main = "QQ plot for smokers")
qqline(bwt.smoke,col=2)

#2. Perform the t test using the following R codes. Compare the resulting p-value with what you
# obtain from the permutation test in Problem 3. Do you reach the same conclusion?

n <- 10
nonsmokers <- sample(bwt.nonsmoke , n)
smokers <- sample(bwt.smoke , n)
result <- t.test(nonsmokers, smokers, alternative = 'two.sided')
result$p.value
```


