---
title: "m254-hw1 Name: Jiahao Tian"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ISLR)
dim(NCI60$data)
head(NCI60$labs) 
table(NCI60$labs)
```
**Problem 1**

1. Order the 6, 830 p-values from the smallest to the largest. Plot the ordered p-values as the y-axis and their corresponding ranks as the x-axis. If all the 6, 830 null hypotheses are true, the p-values should be uniformly distributed. Are the observed p-values likely from a uniform distribution?

**Solution:** Yes, it is uniform distribution.

2. What will be the p-value cutoﬀ if you would like to control the Family Wise Error Rate un-
 der 0.05 using the Bonferroni correction? How many genes will be identiﬁed as differentially
 expressed using this cutoﬀ?
 
3. What will be the p-value cutoﬀ if you would like to control the False Discovery Rate under 0.05
 using the Benjamini-Hochberg procedure? How many genes will be identiﬁed as diﬀerentially
 expressed using this cutoﬀ?
 
4. What will be the p-value cutoﬀ if you would like to control the Per-Comparison Error Rate
 (i.e., signiﬁcance level) under 0.05? How many genes will be identiﬁed as diﬀerentially expressed
 using this cutoﬀ?
 

```{r}
NSCLC <- NCI60$data[NCI60$labs=='NSCLC',] 
RENAL <- NCI60$data[NCI60$labs=='RENAL',] 
p_values <- mapply(function(i) 
  t.test(NSCLC[,i], RENAL[,i], var.equal = T, 
         paired = F, alternative = "two.sided")$p.value, 
  1:ncol(NSCLC))

sort_p <- sort(p_values)
plot(sort_p, xlab = "Rank", ylab = "p value", 
     cex=0.1, col='blue', main = "Sorted p value")

plot(ecdf(sort_p), main="Empirical Dist. of Ordered p-values", col="blue")
lines(x=seq(0,1,0.01), y=seq(0,1,0.01), col="red", lwd=1.5)
text(0.8, 0.6, "Theoretical cdf", col="red")
text(0.25, 0.6, "Empirical cdf", col="blue")

plot(sort_p, cex=0.5, xlim = c(1, 25), ylim = c(0, 0.0008),
     main = "FDR and FWER", xlab = 'index i', ylab = 'p-value', pch=8)

x <- c(1:6830); fdr <- c(x/6830*0.05); bonforroni <- rep(0.05 / 6830, times=length(x));
holm <- c(0.05/(length(x)-x+1))
sum(sort_p <= bonforroni[1])

lines(x, fdr,  lwd=1.5)
lines(x, bonforroni,  lwd=1.5)
lines(x, holm, lwd=0.5)
text(20, 0.00001, "Holm's")
text(23, 0.00001, "Bonforroni")
text(18, 0.0002, "FDR")

 sum(p_values<=0.05)
```
 
**Problem 2**

Use the NCI60 data, give an example to each of the following data analysis tasks (please refer to the Leek and Peng, Science (2015) articl). Note: this is an open question;diﬀerent people may give diﬀ erence examples to the same task.

1. Descriptive study
2. Exploratory study
3. Inferential study
4. Predictive study
5. Causal study
6. Mechanistic study

```{r}
## Descriptive study
table(NCI60$labs)

## Exploratory study
plot(round(apply(NCI60$data, 2, mean)[1:20], 4), xlab="Cell type", 
     ylab="Gene expression", 
     main="Exploratory of first 10 genes", pch=3)

## Inferential study
plot(NCI60$data[,1], NCI60$data[,2], xlab="Gene 1", ylab="Gene 2", cex=0.4, 
     pch=1, main = "Inferential study of gene 1 and 2")

## Predictive study
library(e1071)
dat <- data.frame(NCI60$data[,1:2], y=factor(NCI60$labs))
mod <- svm(y~., data=dat, kernel="linear", cost=10, scale = F)
plot(mod, dat)