---
title: "M254 hw3"
quthor: Jiahao Tian
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("~/Desktop/stat M254/hw3")
library(M3C)
library(acepack)
library(psych)
library(acepack)
library(energy)
library(minerva)
library(XICOR)
```

**Problem 1**

**Solution:**

```{r}
row.names(USArrests)
names(USArrests)
```

**Solution for part 1:**

```{r}
result <- prcomp(USArrests, center = T, scale. = T)
round(result$rotation, 3)
```

**Solution for part 2:**

```{r}
summary(result)
```

**Solution for part 3:**

```{r}
biplot(result, cex=0.6,expand=0.9,col=c('black','red'))
```

From the plot, Maryland, New Mexico, Michigan and Florida form a cluster around murder, 
assault and rape. California and several states form a cluster against urban pop.

The clusters in the plot indicates similarity between samples: for example, in this plot, there's a cluster 
that contains New Jersey, Massachusetts, Hawaii, and Rhode Islands, meaning that these states are highly 
similar to each other. 

The plot shows that UrbanPop have a strong negative impact on both PCs and the angles between vectors 
tell us how characteristics correlate with one another. Small angle between Murder and Assault shows 
that the two variables are positively correlated, and the right angle between Murder and UrbanPop 
shows that the two variables are not likely to be correlated.

**Solution for part 4:**

```{r}
library(magrittr)
library(dplyr)
library(ggpubr)
usas <- USArrests %>%
  dist() %>%          
  cmdscale() %>%
  as_tibble()
colnames(usas) <- c("Dim.1", "Dim.2")
ggscatter(usas, x = "Dim.1", y = "Dim.2", 
          label = rownames(USArrests),
          size = 1, color = "blue",
          repel = TRUE)
```

From the results, t-SNE is good for visualization. The PCA and MDS plots looks not that different.
t-SNE preserves only local similarities, it does not preserve the pairwise distances as MDS, also, does not
pairwise visualized principal compoments as PCA.

**Solution for part 5:**

```{r}
library(M3C)
data = USArrests
umap(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='Murder',])), 
     text = colnames(t(data)))
umap(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='Assault',])), 
     text = colnames(t(data)))
umap(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='UrbanPop',])), 
     text = colnames(t(data)))
umap(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='Rape',])), 
     text = colnames(t(data)))
```

```{r}
library(M3C)
data = USArrests
tsne(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='Murder',])),
     text = colnames(t(data)))
tsne(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='Assault',])), 
     text = colnames(t(data)))
tsne(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='UrbanPop',])), 
     text = colnames(t(data)))
tsne(t(data),
     labels = scale(as.numeric(t(data)[row.names(t(data))=='Rape',])), 
     text = colnames(t(data)))
```

From the results, the scales of Murder, Assault and Rape index looks like some of them are very
high and some of them are very low. But for Urban pop it's like average in every cluster. Also, from the
results, t-SNE is good for visualization. The PCA and MDS plots looks not that different. t-SNE preserves only
local similarities, it does not preserve the pairwise distances as MDS, also, does not
pairwise visualized principal compoments as PCA.

**Problem 2**

**Solution:**

```{r}
B = 1000; n=100

power_H1_pearson <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "pearson"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "pearson"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H1_pearson <- c(power_H1_pearson, 
                 sum(result_alt >= max(threshold) | 
                       result_alt <= min(threshold)) / length(result_alt))
}
print(power_H1_pearson)
```

```{r}
power_H2_pearson <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "pearson"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "pearson"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H2_pearson <- c(power_H2_pearson, 
                 sum(result_alt >= max(threshold) | 
                       result_alt <= min(threshold)) / length(result_alt))
}
print(power_H2_pearson)
```

```{r}
power_H3_pearson <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "pearson"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "pearson"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H3_pearson <- c(power_H3_pearson, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
print(power_H3_pearson)
```

```{r}
power_H4 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "pearson"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "pearson"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H4 <- c(power_H4, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H4_pearson <- power_H4
print(power_H4_pearson)
```

```{r}
power_H5 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(
      n, mean=0, sd=sigma2^0.5
      )
    
    result_null <- c(result_null, cor(x_null, y_null, method = "pearson"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "pearson"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H5 <- c(power_H5, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H5_pearson <- power_H5
print(power_H5_pearson)
```

```{r}
power_H1 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "spearman"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "spearman"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H1 <- c(power_H1, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H1_spearman <- power_H1
print(power_H1_spearman)
```

```{r}
power_H2 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "spearman"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "spearman"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H2 <- c(power_H2, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H2_spearman <- power_H2
print(power_H2_spearman)
```

```{r}
power_H3 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "spearman"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "spearman"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H3 <- c(power_H3, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H3_spearman <- power_H3
print(power_H3_spearman)
```

```{r}
power_H4 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "spearman"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "spearman"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H4 <- c(power_H4, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H4_spearman <- power_H4
print(power_H4_spearman)
```

```{r}
power_H5 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(
      n, mean=0, sd=sigma2^0.5
      )
    
    result_null <- c(result_null, cor(x_null, y_null, method = "spearman"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "spearman"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H5 <- c(power_H5, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H5_spearman <- power_H5
print(power_H5_spearman)
```

```{r}
power_H1 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "kendall"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "kendall"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H1 <- c(power_H1, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H1_kendall<- power_H1
print(power_H1_kendall)
```

```{r}
power_H2 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "kendall"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "kendall"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H2 <- c(power_H2, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H2_kendall <- power_H2
print(power_H2_kendall)
```

```{r}
power_H3 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "kendall"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "kendall"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H3 <- c(power_H3, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H3_kendall <- power_H3
print(power_H3_kendall)
```

```{r}
power_H4 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    result_null <- c(result_null, cor(x_null, y_null, method = "kendall"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "kendall"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H4 <- c(power_H4, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H4_kendall <- power_H4
print(power_H4_kendall)
```

```{r}
power_H5 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(
      n, mean=0, sd=sigma2^0.5
      )
    
    result_null <- c(result_null, cor(x_null, y_null, method = "kendall"))
    result_alt <- c(result_alt, cor(x_alt, y_alt, method = "kendall"))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H5 <- c(power_H5, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H5_kendall <- power_H5
print(power_H5_kendall)
```

```{r}
library(acepack)
power_H1 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
    
    null_ace <- ace(x_null, y_null)
    alt_ace <- ace(x_alt, y_alt)
    result_null <- c(result_null, cor(null_ace$tx, null_ace$ty))
    result_alt <- c(result_alt, cor(alt_ace$tx, alt_ace$ty))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H1 <- c(power_H1, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H1_ace <- power_H1
print(power_H1_ace)
```

```{r}
power_H2 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
    
    null_ace <- ace(x_null, y_null); alt_ace <- ace(x_alt, y_alt)
    result_null <- c(result_null, cor(null_ace$tx, null_ace$ty))
    result_alt <- c(result_alt, cor(alt_ace$tx, alt_ace$ty))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H2 <- c(power_H2, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H2_ace <- power_H2
print(power_H2_ace)
```

```{r}
power_H3 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    null_ace <- ace(x_null, y_null); alt_ace <- ace(x_alt, y_alt)
    result_null <- c(result_null, cor(null_ace$tx, null_ace$ty))
    result_alt <- c(result_alt, cor(alt_ace$tx, alt_ace$ty))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H3 <- c(power_H3, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H3_ace <- power_H3
print(power_H3)
```

```{r}
power_H4 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
    
    null_ace <- ace(x_null, y_null); alt_ace <- ace(x_alt, y_alt)
    result_null <- c(result_null, cor(null_ace$tx, null_ace$ty))
    result_alt <- c(result_alt, cor(alt_ace$tx, alt_ace$ty))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H4 <- c(power_H4, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H4_ace <- power_H4
print(power_H4_ace)
```

```{r}
power_H5 <- c()
for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
  result_null <- c(); result_alt <- c()
  for(i in 1:B) {
    x_null <- rnorm(n); y_null <- rnorm(n)
    x_alt <- rnorm(n); y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(
      n, mean=0, sd=sigma2^0.5
      )
    
    null_ace <- ace(x_null, y_null); alt_ace <- ace(x_alt, y_alt)
    result_null <- c(result_null, cor(null_ace$tx, null_ace$ty))
    result_alt <- c(result_alt, cor(alt_ace$tx, alt_ace$ty))
  }
  threshold <- quantile(result_null, probs=c(0.975, 0.025))
  power_H5 <- c(power_H5, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
}
power_H5_ace <- power_H5
print(power_H5_ace)
```

```{r}
power_dc <- function(alter=1) {
  power_ <- c()
  for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
    result_null <- c(); result_alt <- c()
    for(i in 1:B) {
      x_null <- rnorm(n); y_null <- rnorm(n)
      x_alt <- rnorm(n); 
      if(alter == 5) {
        y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 4) {
        y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 3) {
        y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 2) {
        y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 1) {
        y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
      }
      
      result_null <- c(result_null, DCOR(x_null, y_null, 2)$dCor)
      result_alt <- c(result_alt, DCOR(x_alt, y_alt, 2)$dCor)
    }
    threshold <- quantile(result_null, probs=c(0.975, 0.025))
    power_ <- c(power_, 
                  sum(result_alt >= max(threshold) | 
                        result_alt <= min(threshold)) / length(result_alt))
  }
  power_
}
power_H1_dc <- power_dc(1)
power_H2_dc <- power_dc(2)
power_H3_dc <- power_dc(3)
power_H4_dc <- power_dc(4)
power_H5_dc <- power_dc(5)
```

```{r}
power_mic <- function(alter=1) {
  power_ <- c()
  for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
    result_null <- c(); result_alt <- c()
    for(i in 1:B) {
      x_null <- rnorm(n); y_null <- rnorm(n)
      x_alt <- rnorm(n); 
      if(alter == 5) {
        y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 4) {
        y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 3) {
        y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 2) {
        y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 1) {
        y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
      }
      
      result_null <- c(result_null, cstats(as.matrix(x_null), as.matrix(y_null))[3])
      result_alt <- c(result_alt, cstats(as.matrix(x_alt), as.matrix(y_alt))[3])
    }
    threshold <- quantile(result_null, probs=c(0.975, 0.025))
    power_ <- c(power_, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
  }
  power_
}
power_H1_mic <- power_mic(1)
power_H2_mic <- power_mic(2)
power_H3_mic <- power_mic(3)
power_H4_mic <- power_mic(4)
power_H5_mic <- power_mic(5)
```

```{r}
power_cc <- function(alter=1) {
  power_ <- c()
  for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
    result_null <- c(); result_alt <- c()
    for(i in 1:B) {
      x_null <- rnorm(n); y_null <- rnorm(n)
      x_alt <- rnorm(n); 
      if(alter == 5) {
        y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 4) {
        y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 3) {
        y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 2) {
        y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 1) {
        y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
      }
      
      result_null <- c(result_null, xicor(x_null, y_null))
      result_alt <- c(result_alt, xicor(x_alt, y_alt))
    }
    threshold <- quantile(result_null, probs=c(0.975, 0.025))
    power_ <- c(power_, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
  }
  power_
}
power_H1_cc <- power_cc(1)
power_H2_cc <- power_cc(2)
power_H3_cc <- power_cc(3)
power_H4_cc <- power_cc(4)
power_H5_cc <- power_cc(5)


```

```{r}
library(mvtnorm)
library(gR2)

power_gR2 <- function(alter=1) {
  power_ <- c()
  for(sigma2 in c(.1, .3, .5, .7, .9, 1.1, 1.3, 1.5)) {
    result_null <- c(); result_alt <- c()
    for(i in 1:B) {
      x_null <- rnorm(n); y_null <- rnorm(n)
      x_alt <- rnorm(n); 
      if(alter == 5) {
        y_alt <- x_alt * (-1)**rbinom(1,1,0.5) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 4) {
        y_alt <- sign(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 3) {
        y_alt <- sin(x_alt) + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 2) {
        y_alt <- x_alt ** 2 + rnorm(n, mean=0, sd=sigma2^0.5)
      } else if(alter == 1) {
        y_alt <- x_alt + rnorm(n, mean=0, sd=sigma2^0.5)
      }
      
      result_null <- c(result_null, 
                       gR2(x_null,
                           y_null,
                           K=1,
                           regressionMethod="LM")$estimate)
      result_alt <- c(result_alt, 
                      gR2(x_alt,y_alt,
                          K=1,
                          regressionMethod="LM")$estimate)
    }
    threshold <- quantile(result_null, probs=c(0.975, 0.025))
    power_ <- c(power_, 
                sum(result_alt >= max(threshold) | 
                      result_alt <= min(threshold)) / length(result_alt))
  }
  power_
}
power_H1_gR2 <- power_gR2(1)
power_H2_gR2 <- power_gR2(2)
power_H3_gR2 <- power_gR2(3)
power_H4_gR2 <- power_gR2(4)
power_H5_gR2 <- power_gR2(5)


power <- tibble(power_H1_ace=power_H1_ace,
                    power_H2_ace=power_H2_ace,
                    power_H3_ace=power_H3_ace,
                    power_H4_ace=power_H4_ace,
                    power_H1_spearman = power_H1_spearman,
                    power_H2_spearman = power_H2_spearman,
                    power_H3_spearman = power_H3_spearman,
                    power_H4_spearman = power_H4_spearman,
                    power_H1_pearson = power_H1_pearson,
                    power_H2_pearson = power_H2_pearson,
                    power_H3_pearson = power_H3_pearson,
                    power_H4_pearson = power_H4_pearson,
                    power_H1_kendall = power_H1_kendall,
                    power_H2_kendall = power_H2_kendall,
                    power_H3_kendall = power_H3_kendall,
                    power_H4_kendall = power_H4_kendall,
                    power_H1_dc = power_H1_dc,
                    power_H2_dc = power_H2_dc,
                    power_H3_dc = power_H3_dc,
                    power_H4_dc = power_H4_dc,
                    power_H1_mic = power_H1_mic,
                    power_H2_mic = power_H2_mic,
                    power_H3_mic = power_H3_mic,
                    power_H4_mic = power_H4_mic,
                    power_H1_cc = power_H1_cc,
                    power_H2_cc = power_H2_cc,
                    power_H3_cc = power_H3_cc,
                    power_H4_cc = power_H4_cc,
                    power_H1_gR2 = power_H1_gR2, 
                    power_H2_gR2 = power_H2_gR2,
                    power_H3_gR2,
                    power_H4_gR2)
write.csv(power, 'power.csv')
power
```

Conclusion:
Distance correlation has power estimates across all tests at all sigma square values.Pearson and spearman rank
correlations have higher power estimates on all tests except H0 vs H2. they do not work well in square relationship 
or nonlinear relationship.

Generalized Pearson correlation squares has fluctuated power estimates on H0 vs H4. In general the value being 
very high. Maximal correlation and maximal information coefficient has decreasing power estimates in 
case H0 vs H3 and H0 vs H1, when sigma square increases shown in the graph. For H0 vs H1, the decreasing scale 
is minor to observe.

Kendall is not shown clearly in the graph but according to the value we get it is similar to pearson and spearman.

Chatterjee has a drop in H0 vs H1 and H0 vs H3, and it also has minor drop in H2 and H4.

**Problem 3**

**Solution:**

```{r}
library(psych)
library(psychTools)
library(XICOR)
library(energy)
library(minerva)
library(gR2)
data(peas)
cor(peas$parent, peas$child, method = 'pearson')
xicor(peas$parent, peas$child)
xicor(peas$child, peas$parent)
table = table(peas$child, peas$parent)
table
cor(peas$parent, peas$child, method = 'pearson')
cor(peas$parent, peas$child, method = 'spearman')
cor(peas$parent, peas$child, method = 'kendall')
ace(peas$parent, peas$child)$rsq
dcor(peas$parent, peas$child)
mine(peas$parent, peas$child)$MIC
xicor(peas$parent, peas$child)
gR2(peas$parent, peas$child, K=2, mc.cores = 1)$estimate
cor(peas$child, peas$parent, method = 'pearson')
cor(peas$child, peas$parent, method = 'spearman')
cor(peas$child, peas$parent, method = 'kendall')
ace(peas$child, peas$parent)$rsq
dcor(peas$child, peas$parent)
mine(peas$child, peas$parent)$MIC
xicor(peas$child, peas$parent)
gR2(peas$child, peas$parent, K=2, mc.cores = 1)$estimate
```

From the results, when change the position of child and parent, all functions gives correlation
coefficient number except xicor function. Therefore, when compare to other measurements, xicor
took nonlinear dependency into account.
